---
title: "script"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(jsonlite)
library(scales)

responses <- fromJSON("../data/form_responses.json") %>% 
  mutate(
    income = factor(income, levels = c("Less than $40K", "$40K to $80K", "$80K to $125K", "$125K to $250K", "$250K to $350K", "$350K to $500K", "Greater than $500K")),
    net_price = factor(net_price, levels= c("0", "Less than $5K", "$10K to $15K", "$15K to $20K", "$20K to $25K", "$25K to $30K", "$30K to $35K", "$35K to $40K", "$45K to $50K", "$50K to $55K", "$55K to $60K", "$60K to $65K", "$65K to $70K", "$70K to $75K", "More than $75K")),
)

responses_race <- responses %>% 
  separate(categories, sep = "; ", into = c("category1", "category2", "category3")) %>% 
  gather(category_num, category, category1:category3) %>% 
  filter(!is.na(category))

responses_ec <- responses %>% 
  separate(extracurriculars, sep = ", ", into = c("ec1", "ec2", "ec3", "ec4", "ec5", "ec6", "ec7", "ec8")) %>% 
  gather(extracurricular_num, extracurricular, ec1:ec8) %>% 
  filter(!is.na(extracurricular)) %>% 
  group_by(extracurricular) %>% 
  filter(n() > 10) %>% 
  ungroup()
responses_race_ec <- responses_ec %>% 
  separate(categories, sep = "; ", into = c("category1", "category2", "category3")) %>% 
  gather(category_num, category, category1:category3) %>% 
  filter(!is.na(category))

responses_q <- responses %>%
  select(starts_with("q_")) %>% 
  filter(q_fossil_fuels_divest != "")
```

## Income and race

Income distribution by race
```{r}
responses_race %>%
  ggplot(aes(category, fill = income)) +
  geom_bar(position = "fill") +
  coord_flip()
```

## Extracurriculars

Conservatives slightly more likely to join political organizations and student government, liberals much more likely to join social justice/activism and culture or identity groups.
```{r}
responses_ec %>%
  filter(!political_af %in% c("I don't know", "Neither conservative nor liberal")) %>% 
  filter(!school %in% c("Barnard College", "General Studies")) %>%  # barnard not enough conservatives, gs not enough students
  filter(income %in% c("Greater than $500K", "$125K to $250K", "Less than $40K")) %>% # only enough conservatives in these buckets
  mutate(
    political_af = as.character(political_af),
    political_af = if_else(political_af == "Very conservative", "Somewhat conservative", political_af),
    political_af = if_else(political_af == "Very liberal", "Somewhat liberal", political_af),
  ) %>% 
  ggplot(aes(income, fill = extracurricular)) +
  geom_bar(position = "fill") + 
  coord_flip() +
  facet_grid(~ political_af, scales = "free")
```

Asian
- Richer = more sports, ~more visual and performing arts, ~less student government, ~less political organizations, ~more campus pubs
- No matter what income, same for: culture/identity groups,

Black
- Richer = more sports
- No matter what income, same for: culture/identity groups

Hispanic/Latino:
- Richer = ~more campus publications, culture/identity interest goes down, more visual/performing arts

White:
- No matter what come, same for: Campus pubs, sports, political orgs, social justice, less student gov't, visuall/performing arts

```{r}
responses_race_ec %>%
  filter(!category %in% c("Native Hawaiian or Other Pacific Islander")) %>% 
  # mutate(
    # extracurricular = if_else(extracurricular %in% c("Culture or Identity Groups", "Political Organizations", "Social Justice/Activism"), extracurricular, "Other")
  # ) %>% 
  ggplot(aes(income, fill = extracurricular)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~.)
```

## Socioeconomic diversity

Asians and Whites are spread out pretty evenly in all income brackets, but Blacks and Hispanics are skewed to low-income.

**More poor students are first generation.**
```{r}
responses_race %>% 
  filter(!category %in% c("Native Hawaiian or Other Pacific Islander")) %>% 
  mutate(
    category = factor(category, levels = c("White", "Asian", "Hispanic, Latino, or Spanish origin", "Black or African American", "American Indian or Alaskan Native", "Middle Eastern or North African"))
  ) %>% 
  ggplot(aes(income)) +
  geom_bar() +
  facet_grid(~ category) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
```

Support for affirmative action among Asians and whites peaks don't show too strongly of a trend , but it seems that higher-income students tend to not support affirmative action. But
```{r}
responses %>% 
  group_by(income, category) %>% 
  filter(n() > 10) %>% 
  ungroup() %>% 
  ggplot(aes(income, fill = affirmative_action)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ .)
```


## Voting

```{r}
responses %>% 
  filter(vote_primary != "" && vote_gubernatorial != "" && vote_senate != "" && vote_house != "") %>% 
  filter(!political_af %in% c("I don't know", "Neither conservative nor liberal")) %>% 
  mutate(
    political_af = as.character(political_af),
    political_af = if_else(political_af == "Very conservative", "Somewhat conservative", political_af),
    political_af = if_else(political_af == "Very liberal", "Somewhat liberal", political_af),
  ) %>% 
  ggplot(aes(income, fill = vote_primary)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ .)
```

## College decisions factors

```{r}
decisions <- responses %>% 
  gather(decision_factor, decision_priority, starts_with("decision"))

decisions %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  # filter(hs_type != "Homeschooled") %>% 
  ggplot(aes(income, fill = decision_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ decision_factor)
```

## Columbia factors



```{r}
factors <- responses %>% 
  gather(factor, factor_priority, starts_with("factor_"))

factors %>% 
  filter(!factor %in% c("factor_cost", "factor_resources", "factor_core", "factor_abroad", "factor_research", "factor_extracurriculras", "factor_athletics")) %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  mutate(
    factor_priority = if_else(factor_priority == "Strongly negative", "Negative", factor_priority),
    factor_priority = if_else(factor_priority == "Strongly positive", "Positive", factor_priority),
    factor_priority = factor(factor_priority, levels = c("Negative", "Indifferent", "Positive"))
  ) %>% 
  ggplot(aes(income, fill = factor_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ factor)
```


```{r}
factors <- responses %>% 
  gather(factor, factor_priority, starts_with("factor_"))

factors %>% 
  filter(factor %in% c("factor_cost", "factor_resources",
                       #"factor_core", we have a question about this already
                       "factor_demographics",
                       "factor_greek",
                       #"factor_mental_health", we have a question about this already
                       "factor_social")) %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  mutate(
    factor_priority = if_else(factor_priority == "Strongly negative", "Negative", factor_priority),
    factor_priority = if_else(factor_priority == "Strongly positive", "Positive", factor_priority),
    factor_priority = factor(factor_priority, levels = c("Negative", "Indifferent", "Positive"))
  ) %>% 
  ggplot(aes(income, fill = factor_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ factor)
```


